Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 egpaper_final.tex
minor changes
breaking lines

%     Using 3D model to detect and register the model to RGB image has been a growing field of study. Recently, Aubry \etal \cite{Aubry14} and J. Lim \etal \cite{Lim14} tried to estimate pose and align exact CAD model to an image using part-based templates. Malisiewicz \etal \cite{Malisiewicz11} also addresses the issue by transferring metadata after detection. Hejrati \etal \cite{Hejrati14} uses template synthesis to detect an object. All these approaches require extensive training, a lot of memory space. Also, the training based approach requires defining fixed viewpoints which limit themselves.
%  To overcome these difficulties, we propose an efficient and fast way to synthesize and validate templates using on the fly realistic rendering that can jointly optimize scale, translation, rotation and focal length continuously. We render an object and extract features and decorrelate spatial dependencies within them on the fly to make a discriminative template. To speed up the decorrelation procedure, we adopt a gradient based algorithm which is numerically more stable than decomposition procedure. Finally, we efficiently search for candidate match by computing convolution using FFT and jointly optimize to match the object accurately. To speed up further, we propose an efficient GPU implementation and tested on PASCAL images.

%   In contrast to other mid-level patch based 3D matching methods which requires extensive training and more than 10Gb of memory space and physical storage, ours require no training nor saving the templates thus enabling it to run on any personal computer. 
%   To overcome these difficulties, we propose an efficient and fast way to synthesize and validate templates using on the fly realistic rendering that can jointly optimize scale, translation, rotation and focal length continuously. We render an object and extract features and decorrelate spatial dependencies within them on the fly to make a discriminative template. To speed up the decorrelation procedure, we adopt a gradient based algorithm which is numerically more stable than decomposition procedure. Finally, we efficiently search for candidate match by computing convolution using FFT and jointly optimize to match the object accurately. To speed up further, we propose an efficient GPU implementation and tested on PASCAL images.

@@

% specialized in giving such accurate information, can be applied on any generic object detectors to give high quality metadata.
% Many 
% Since the WHO does not require extensive training such as \cite{Felzenszwalb10, Malisiewicz11, Girshick14}, WHO is gaining momentum in 2D-3D matching. \cite{Aubry13, Aubry14, Lim14} combined WHO with synthetic rendering to jointly detect and estimate pose by making mid-level patches using WHO. Using this
% WSuch approach has an advantage over generic detections. First, it is easy to trace the detection to specific template. The template also has The rendering has corresponding CAD model and viewpoint thus enabling joint detection and viewpoint estimation. 
% To deal with textureless background \cite{Aubry13} whitens HOG feature and then zero out the WHO feature where there is no texture. But centering a HOG feature where there is no edge creates strong negative edges. We found out that centering only non-zero cells help detection but whitening all cell makes features to leak out to texture-less regions.
% With increasing popularity of rendering image, whitening, traditional signal processing technique, as a preprocessing stage has been widely used also in computer vision community.
% Many object detectors \cite{Felzenszwalb10, Malisiewicz11}
% \subsection{Mid-level Patches for 2D-3D matching}
%  FPM introduction : per instance matching  
%  Seeing 3D chair : mid-level patches

@@

We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), that are both more discriminative than WHO, and can be generated in 70ms (compared to several seconds for WHO). This speed up means we can generate templates on the fly, allowing our approach to evaluate arbitrary viewpoint hypotheses. 
\scream{chris:We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), an improvement over WHO which is both more discriminative than WHO and can be computed many orders of magnitude faster than WHO (70ms as opposed to to several seconds for WHO). This speedup allows us to generate NZ-WHO templates on the fly, allowing us to quickly evaluate arbitrary viewpoint hypotheses.}

The computational efficiency of our approach comes from using the iterative conjugate gradient method to whiten the HOG template. Whitening a feature vector $x$ requires a system of linear equations $\Sigma w = (x - \mu)$ to be solved, usually by decomposing the covariance matrix $\Sigma$ (with LU or Cholesky decomposition). Since the covariance matrices of HOG templates have nice spectral properties (see Section \ref{sec:fast_whitening}), the conjugate gradient method will converge to the solution several orders of magnitude faster than decomposition methods.

@@

% To accommodate the growing usage of rendering image for 2D-3D matching using WHO\cite{Aubry13, Aubry14, Lim14}, we analyzed various methods to whiten synthetic rendering image and propose Non-Zero Whitened Histograms of Orientations (NZ-WHO) which is a direct extension of \cite{Hariharan12} designed for rendering images with textureless background. % Since a NZ-WHO template has zero-mean and all elements in the template follow same distribution, no explicit calibration is required if all the templates have approximately the same number of cells.

@@

% Since the statistics have been computed for a generic HOG feature, we can use the same statistics for all templates and images. In practice, since it is expensive to whiten all the sliding patches in the image, we followed \cite{Hariharan12} and defined the templates to be $w = \Sigma^{-1}(X-E[X])$\begin{comment}which is a good approximation of $\Sigma^{-\frac{T}{2}} \Sigma^{-\frac{1}{2}}(X-E[X])$\end{comment}

% \begin{comment}Gathering statistics, especially computing spatial covariance of a signal is computationally burdensome. \end{comment}
% To speed up the statistics collection, \begin{comment}in addition to the wide-sense stationarity (WSS) of HOG, \end{comment}we assumed horizontal and vertical symmetry of covariance.

 %Hariharan \etal \cite{Hariharan12} interpreted the whitening also as the Linear Discriminant Analysis where each class has the same covariance $\Sigma$ but with different mean value. They defined this whitened HOG feature as Whitened Histograms of Orientations (WHO) and we will follow their convention of the name.

% \subsection{Rendering}
% 
% We used a public rendering engine to create realistic rendering and depth\begin{comment}\cite{Choy14render}\end{comment}. The CAD models we used in our experiments contain color, texture and material information such as transparency and reflectance. We tried to make the rendering as much realistic as possible to simulate the natural image statistics. The example rendering is in Fig. \ref{fig:rendering}. We also extracted depth information as well while rendering to use in the reconstruction.

% To whiten a HOG features, we first have to synthesize the covariance matrix. Following the method of \cite{Hariharan12}, which assumes spatial Wide-Sense Stationarity (WSS) of a feature, covariance matrix can be synthesized using spatial autocovariance $\Gamma$. This assumes the \begin{comment} and features such as HOG are gradients in a small patch whose statistics does not change (an object can be in anywhere in an image)\end{comment}.

%However, unlike small fixed-sized mid-level patches, the root templates have different aspect ratios, different gradient patterns and large number of HOG cells which results in large covariance $\Sigma \in \mathcal{R}^{7000 \times 7000}$ that is difficult to cache. Thus, to generate a NZ-WHO template, we have to synthesize a covariance matrix and solve the system of linear equations.

% Thus, every time we whiten a root template, we have to generate a covariance matrix using autocovariance $\Gamma$.

%\begin{comment} Also, since we are whitening only the non-zero HOG cells whose position is unknown beforehand and all templates have different aspect ratio for , we cannot cache the matrix. Using CPU, it takes about \end{comment}

% In practice, synthesizing covariance matrix $\Sigma$ takes long time due to its large size.  $\Sigma$ ranges from $\Sigma \in \mathcal{R}^{7000 \times 7000}$ to $\Sigma \in \mathcal{R}^{7500 \times 7500}$. We vary the template resolution and run $\Sigma$ generation on CPU and GPU 100 times and report the average time in Fig. \ref{fig:covariancetime}.

@@

% where $w$ is the weight we want to find and $x$ is the HOG feature from the rendering and $\lambda$ is the regularization constant (see section \ref{sec:regularization}). For $w\in \mathcal{R}^n$, it takes $O(n^3)$ time to solve the linear equation. If we use a high resolution template to capture small geometric properties we can get high accuracy matchings but the solving the linear equation becomes computationally burdensome.

% To briefly recap, the LU Decomposition and Cholesky Decomposition are most popular methods to solve a linear equation. In linear algebra libraries such as MATLAB, these decomposition methods are default method to solve a linear equation (backslash \textbackslash in MATLAB). Since the covariance matrix is Hermitian and Positive Semi-Definite, we can use Cholesky Decomposition which is twice as fast as the LU Decomposition.

% decomposition methods such as LU or Cholesky Decomposition not only take long time but they also accumulate numerical error.

%In our setting, we used high resolution template which has more than 250 HOG cells. This result in $\Sigma \in \mathcal{R}^{7500 \times 7500}$. It takes about 9 seconds if we use LU decomposition and 5 seconds if we use Cholesky decomposition.  % The default way of solving the linear equation in many linear algebra library is decomposing the matrix using LU Decomposition (in MATLAB the default method of backslash operator) which takes $O(\frac{2}{3}n^3)$ flops for $n\times n$ matrix. Since the matrix, $\Sigma + \lambda I$ is Hermitian and Positive Semi Definite. We can speed up using Cholesky Decomposition.

%We also used iterative Conjugate Gradient to make use of low rank property of the covariance \cite{Gharbi12}. After adding small regularization, the regularized covariance matrix has nice spectral property (Figure \ref{fig:spectral}). The singular values are clustered in small region and most of the energies are concentrated on the first (number) singular values. Thus, the conjugate gradients stage doesn't go upto the size of the matrix to converge.

@@

% \subsection{Effect of Regularization}
% \label{sec:regularization}
%  Also, since all covariance matrix is Positive Semi Definite, our matrix $\Sigma$ must be Positive Semi-Definite too. To guarantee the property, We found out that the regularization has significant impact on the performance. In \cite{Hariharan12}, Hariharan \etal used small fixed regularization since the covariance matrix of specially correlated features is intrinsically row-rank to make the matrix invertible.
%  
% However if we closely analyze the impact of the regularization, we can see that larger the $\lambda$ is, the centered HOG becomes more dominant. 
% 
% \begin{align}
% (\Sigma + \lambda I)^{-1} & = \frac{1}{\lambda}\left(I + \frac{1}{\lambda}\Sigma \right)^{-1} \\
% & = \frac{1}{\lambda}\left(I - \frac{1}{\lambda} \Sigma  + \frac{1}{\lambda^2} \Sigma^{2} 
% 		- \frac{1}{\lambda^3} \Sigma^{3} + \cdots \right)
% % & = \frac{1}{\lambda}I - \left(I + \frac{1}{\lambda} \Sigma \right)^{-1} \frac{1}{\lambda^2} \Sigma
% % & = \frac{1}{\lambda}\left(I  - \frac{1}{1 + \frac{1}{\Sigma} \Tr{\Sigma}} \right)
% \end{align} 
% 
% We empirically found out that as $\lambda$ increases, AP converges to that of centered HOG feature without any whitening.

@@

In this section, we validate our proposed NZ-WHO templates in two different ways. First, we make a set of NZ-WHO templates for a set of CAD models from various viewpoints. Then we run the bank of NZ-WHO templates as ensembles of weak detectors and show that the rendering image can be We want to see how robust the NZ-WHO to real images a

>>>>

%     Using 3D model to detect and register the model to RGB image has been a
%     growing field of study. Recently, Aubry \etal \cite{Aubry14} and J. Lim
%     \etal \cite{Lim14} tried to estimate pose and align exact CAD model to an
%     image using part-based templates. Malisiewicz \etal \cite{Malisiewicz11}
%     also addresses the issue by transferring metadata after detection. Hejrati
%     \etal \cite{Hejrati14} uses template synthesis to detect an object. All
%     these approaches require extensive training, a lot of memory space. Also,
%     the training based approach requires defining fixed viewpoints which limit
%     themselves.  To overcome these difficulties, we propose an efficient and
%     fast way to synthesize and validate templates using on the fly realistic
%     rendering that can jointly optimize scale, translation, rotation and focal
%     length continuously. We render an object and extract features and
%     decorrelate spatial dependencies within them on the fly to make a
%     discriminative template. To speed up the decorrelation procedure, we adopt
%     a gradient based algorithm which is numerically more stable than
%     decomposition procedure. Finally, we efficiently search for candidate
%     match by computing convolution using FFT and jointly optimize to match the
%     object accurately. To speed up further, we propose an efficient GPU
%     implementation and tested on PASCAL images.

%   In contrast to other mid-level patch based 3D matching methods which
%   requires extensive training and more than 10Gb of memory space and physical
%   storage, ours require no training nor saving the templates thus enabling it
%   to run on any personal computer.  To overcome these difficulties, we propose
%   an efficient and fast way to synthesize and validate templates using on the
%   fly realistic rendering that can jointly optimize scale, translation,
%   rotation and focal length continuously. We render an object and extract
%   features and decorrelate spatial dependencies within them on the fly to make
%   a discriminative template. To speed up the decorrelation procedure, we adopt
%   a gradient based algorithm which is numerically more stable than
%   decomposition procedure. Finally, we efficiently search for candidate match
%   by computing convolution using FFT and jointly optimize to match the object
%   accurately. To speed up further, we propose an efficient GPU implementation
%   and tested on PASCAL images.

@@

% specialized in giving such accurate information, can be applied on any generic
% object detectors to give high quality metadata.  Many Since the WHO does not
% require extensive training such as \cite{Felzenszwalb10, Malisiewicz11,
% Girshick14}, WHO is gaining momentum in 2D-3D matching. \cite{Aubry13,
% Aubry14, Lim14} combined WHO with synthetic rendering to jointly detect and
% estimate pose by making mid-level patches using WHO. Using this WSuch approach
% has an advantage over generic detections. First, it is easy to trace the
% detection to specific template. The template also has The rendering has
% corresponding CAD model and viewpoint thus enabling joint detection and
% viewpoint estimation.  To deal with textureless background \cite{Aubry13}
% whitens HOG feature and then zero out the WHO feature where there is no
% texture. But centering a HOG feature where there is no edge creates strong
% negative edges. We found out that centering only non-zero cells help detection
% but whitening all cell makes features to leak out to texture-less regions.
% With increasing popularity of rendering image, whitening, traditional signal
% processing technique, as a preprocessing stage has been widely used also in
% computer vision community.  Many object detectors \cite{Felzenszwalb10,
% Malisiewicz11} \subsection{Mid-level Patches for 2D-3D matching} FPM
% introduction : per instance matching  Seeing 3D chair : mid-level patches

@@

We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), that are both
more discriminative than WHO, and can be generated in 70ms (compared to several
seconds for WHO). This speed up means we can generate templates on the fly,
allowing our approach to evaluate arbitrary viewpoint hypotheses.

\scream{chris:We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO),
an improvement over WHO which is both more discriminative than WHO and can be
computed many orders of magnitude faster than WHO (70ms as opposed to to several
seconds for WHO). This speedup allows us to generate NZ-WHO templates on the
fly, allowing us to quickly evaluate arbitrary viewpoint hypotheses.}

The computational efficiency and numerical stability of our approach comes from
using the iterative conjugate gradient method to whiten the HOG template.
Whitening a feature vector $x$ requires a system of linear equations $\Sigma w =
(x - \mu)$ to be solved, usually by decomposing the covariance matrix $\Sigma$
(with LU or Cholesky decomposition). Since the covariance matrices of HOG
templates have nice spectral properties (see Section \ref{sec:fast_whitening}),
the conjugate gradient method will converge to the solution several orders of
magnitude faster than decomposition methods.

@@

% To accommodate the growing usage of rendering image for 2D-3D matching using
% WHO\cite{Aubry13, Aubry14, Lim14}, we analyzed various methods to whiten
% synthetic rendering image and propose Non-Zero Whitened Histograms of
% Orientations (NZ-WHO) which is a direct extension of \cite{Hariharan12}
% designed for rendering images with textureless background. % Since a NZ-WHO
% template has zero-mean and all elements in the template follow same
% distribution, no explicit calibration is required if all the templates have
% approximately the same number of cells.

@@

% Since the statistics have been computed for a generic HOG feature, we can use
% the same statistics for all templates and images. In practice, since it is
% expensive to whiten all the sliding patches in the image, we followed
% \cite{Hariharan12} and defined the templates to be $w =
% \Sigma^{-1}(X-E[X])$\begin{comment}which is a good approximation of
% $\Sigma^{-\frac{T}{2}} \Sigma^{-\frac{1}{2}}(X-E[X])$\end{comment}

% \begin{comment}Gathering statistics, especially computing spatial covariance
% of a signal is computationally burdensome. \end{comment} To speed up the
% statistics collection, \begin{comment}in addition to the wide-sense
% stationarity (WSS) of HOG, \end{comment}we assumed horizontal and vertical
% symmetry of covariance.

 %Hariharan \etal \cite{Hariharan12} interpreted the whitening also as the
 %Linear Discriminant Analysis where each class has the same covariance $\Sigma$
 %but with different mean value. They defined this whitened HOG feature as
 %Whitened Histograms of Orientations (WHO) and we will follow their convention
 %of the name.

% \subsection{Rendering}
% 
% We used a public rendering engine to create realistic rendering and
% depth\begin{comment}\cite{Choy14render}\end{comment}. The CAD models we used
% in our experiments contain color, texture and material information such as
% transparency and reflectance. We tried to make the rendering as much realistic
% as possible to simulate the natural image statistics. The example rendering is
% in Fig. \ref{fig:rendering}. We also extracted depth information as well while
% rendering to use in the reconstruction.

% To whiten a HOG features, we first have to synthesize the covariance matrix.
% Following the method of \cite{Hariharan12}, which assumes spatial Wide-Sense
% Stationarity (WSS) of a feature, covariance matrix can be synthesized using
% spatial autocovariance $\Gamma$. This assumes the \begin{comment} and features
% such as HOG are gradients in a small patch whose statistics does not change
% (an object can be in anywhere in an image)\end{comment}.

%However, unlike small fixed-sized mid-level patches, the root templates have
%different aspect ratios, different gradient patterns and large number of HOG
%cells which results in large covariance $\Sigma \in \mathcal{R}^{7000 \times
%7000}$ that is difficult to cache. Thus, to generate a NZ-WHO template, we have
%to synthesize a covariance matrix and solve the system of linear equations.

% Thus, every time we whiten a root template, we have to generate a covariance
% matrix using autocovariance $\Gamma$.

%\begin{comment} Also, since we are whitening only the non-zero HOG cells whose
%position is unknown beforehand and all templates have different aspect ratio
%for , we cannot cache the matrix. Using CPU, it takes about \end{comment}

% In practice, synthesizing covariance matrix $\Sigma$ takes long time due to
% its large size.  $\Sigma$ ranges from $\Sigma \in \mathcal{R}^{7000 \times
% 7000}$ to $\Sigma \in \mathcal{R}^{7500 \times 7500}$. We vary the template
% resolution and run $\Sigma$ generation on CPU and GPU 100 times and report the
% average time in Fig. \ref{fig:covariancetime}.

@@

% where $w$ is the weight we want to find and $x$ is the HOG feature from the
% rendering and $\lambda$ is the regularization constant (see section
% \ref{sec:regularization}). For $w\in \mathcal{R}^n$, it takes $O(n^3)$ time to
% solve the linear equation. If we use a high resolution template to capture
% small geometric properties we can get high accuracy matchings but the solving
% the linear equation becomes computationally burdensome.

% To briefly recap, the LU Decomposition and Cholesky Decomposition are most
% popular methods to solve a linear equation. In linear algebra libraries such
% as MATLAB, these decomposition methods are default method to solve a linear
% equation (backslash \textbackslash in MATLAB). Since the covariance matrix is
% Hermitian and Positive Semi-Definite, we can use Cholesky Decomposition which
% is twice as fast as the LU Decomposition.

% decomposition methods such as LU or Cholesky Decomposition not only take long
% time but they also accumulate numerical error.

%In our setting, we used high resolution template which has more than 250 HOG
%cells. This result in $\Sigma \in \mathcal{R}^{7500 \times 7500}$. It takes
%about 9 seconds if we use LU decomposition and 5 seconds if we use Cholesky
%decomposition.  % The default way of solving the linear equation in many linear
%algebra library is decomposing the matrix using LU Decomposition (in MATLAB the
%default method of backslash operator) which takes $O(\frac{2}{3}n^3)$ flops for
%$n\times n$ matrix. Since the matrix, $\Sigma + \lambda I$ is Hermitian and
%Positive Semi Definite. We can speed up using Cholesky Decomposition.

%We also used iterative Conjugate Gradient to make use of low rank property of
%the covariance \cite{Gharbi12}. After adding small regularization, the
%regularized covariance matrix has nice spectral property (Figure
%\ref{fig:spectral}). The singular values are clustered in small region and most
%of the energies are concentrated on the first (number) singular values. Thus,
%the conjugate gradients stage doesn't go upto the size of the matrix to
%converge.

@@

% \subsection{Effect of Regularization}
% \label{sec:regularization}
%  Also, since all covariance matrix is Positive Semi Definite, our matrix
%  $\Sigma$ must be Positive Semi-Definite too. To guarantee the property, We
%  found out that the regularization has significant impact on the performance.
%  In \cite{Hariharan12}, Hariharan \etal used small fixed regularization since
%  the covariance matrix of specially correlated features is intrinsically
%  row-rank to make the matrix invertible.
%  
% However if we closely analyze the impact of the regularization, we can see that larger the $\lambda$ is, the centered HOG becomes more dominant. 
% 
% \begin{align}
% (\Sigma + \lambda I)^{-1} & = \frac{1}{\lambda}\left(I + \frac{1}{\lambda}\Sigma \right)^{-1} \\
% & = \frac{1}{\lambda}\left(I - \frac{1}{\lambda} \Sigma  + \frac{1}{\lambda^2} \Sigma^{2} 
% 		- \frac{1}{\lambda^3} \Sigma^{3} + \cdots \right)
% % & = \frac{1}{\lambda}I - \left(I + \frac{1}{\lambda} \Sigma \right)^{-1} \frac{1}{\lambda^2} \Sigma
% % & = \frac{1}{\lambda}\left(I  - \frac{1}{1 + \frac{1}{\Sigma} \Tr{\Sigma}} \right)
% \end{align} 
% 
% We empirically found out that as $\lambda$ increases, AP converges to that of centered HOG feature without any whitening.

@@

In this section, we validate our proposed NZ-WHO templates in two different ways. First, we make a set of NZ-WHO templates for a set of CAD models from various viewpoints. Then we run the bank of NZ-WHO templates as an ensemble of weak detectors and show that we can achieve 2D-3D matching We want to see how robust the NZ-WHO to real images
